SPARQL on the Open, Decentralised Web

### 2016-08-03 dissertation/introduction.tex
Make examples fit #18

\begin{example}{Turtle syntax}\label{ex:turtle}
\begin{verbatim}
@prefix foaf: <http://xmlns.com/foaf/0.1/>.
@prefix dct:  <http://purl.org/dc/terms/>.

<http://www.kjetil.kjernsmo.net/foaf#me> a foaf:Person ;
    foaf:name "Kjetil Kjernsmo" ;
    foaf:made <http://folk.uio.no/kjekje/2013/iswc.pdf> .
<http://folk.uio.no/kjekje/2013/iswc.pdf> a foaf:Document ;
    dct:title "Introducing Statistical Design of Experiments to SPARQL Endpoint Evaluation" ; 
    dct:creator <http://www.kjetil.kjernsmo.net/foaf#me>, 
                <http://data.semanticweb.org/person/john-s-tyssedal> .
\end{verbatim}
\end{example}

The example opens with prefix declarations, to allow abbreviating
IRIs, e.g. the predicate \rdfterm{foaf:name} is an abbreviation of the
IRI \rdfterm{\nolinkurl{http://xmlns.com/foaf/0.1/name}}. The IRI
\rdfterm{\nolinkurl{http://www.kjetil.kjernsmo.net/foaf\#me}} is a name for this
author. The \rdfterm{a} predicate is standards-defined abbreviation
for \rdfterm{\nolinkurl{http://www.w3.org/1999/02/22-rdf-syntax-ns\#type}}, which
is used for declaring the subject an instance of a class, in this case
\rdfterm{foaf:Person}. Then, the \rdfterm{foaf:made} predicate
is used to link to an author-published version of one of the
publications of this dissertation, and further information is then
provided about this publication, including a link to the co-author. As
we see, semicolons are used to separate statements that share a
subject, commas separate statements with the same subject and
predicate, while periods terminate a statement.

@@

\begin{example}{A triple pattern}
\begin{verbatim}
?subject a ?class .
\end{verbatim}
\end{example}
This triple pattern will match the two triples 
\begin{example}{Triples matched by triple pattern}
\begin{verbatim}
<http://www.kjetil.kjernsmo.net/foaf#me> a foaf:Person .
<http://folk.uio.no/kjekje/2013/iswc.pdf> a foaf:Document .
\end{verbatim}
\end{example}
from the above example (prefixes omitted for brevity). Frameworks in
popular programming languages commonly have a method that implements
this.

>>>>

\begin{example}{Turtle syntax}\label{ex:turtle}
\small
\begin{verbatim}
@prefix foaf: <http://xmlns.com/foaf/0.1/>.
@prefix dct:  <http://purl.org/dc/terms/>.

<http://www.kjetil.kjernsmo.net/foaf#me> a foaf:Person ;
  foaf:name "Kjetil Kjernsmo" ;
  foaf:made <http://folk.uio.no/kjekje/2013/iswc.pdf> .
<http://folk.uio.no/kjekje/2013/iswc.pdf> a foaf:Document ;
  dct:title "Introducing Statistical Design of [...]" ; 
  dct:creator <http://www.kjetil.kjernsmo.net/foaf#me>, 
              <http://[...]/person/john-s-tyssedal> .
\end{verbatim}
\normalsize
\end{example}

The example opens with prefix declarations, to allow abbreviating
IRIs, e.g. the predicate \rdfterm{foaf:name} is an abbreviation of the
IRI \rdfterm{\nolinkurl{http://xmlns.com/foaf/0.1/name}}. The IRI
\rdfterm{\nolinkurl{http://www.kjetil.kjernsmo.net/foaf\#me}} is a name for this
author. The \rdfterm{a} predicate is standards-defined abbreviation
for \rdfterm{\nolinkurl{http://www.w3.org/1999/02/22-rdf-syntax-ns\#type}}, which
is used for declaring the subject an instance of a class, in this case
\rdfterm{foaf:Person}. Then, the \rdfterm{foaf:made} predicate
is used to link to an author-published version of one of the
publications of this dissertation, and further information is then
provided about this publication, including a link to the co-author. As
we see, semicolons are used to separate statements that share a
subject, commas separate statements with the same subject and
predicate, while periods terminate a statement. \texttt{[...]} is just
to shorten the lines in the example and is not a part of the syntax.

@@

\begin{example}{A triple pattern}
\begin{verbatim}
?subject a ?class .
\end{verbatim}
\end{example}
This triple pattern will match the two triples 
\begin{example}{Triples matched by triple pattern}
\small
\begin{verbatim}
<http://www.kjetil.kjernsmo.net/foaf#me> a foaf:Person .
<http://folk.uio.no/kjekje/2013/iswc.pdf> a foaf:Document .
\end{verbatim}
\normalsize
\end{example}
from the above example (prefixes omitted for brevity). Frameworks in
popular programming languages commonly have a method that implements
this.
